---
title: "FOIA Data Analysis_v5"
author: "P. Zacher"
date: "2023-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(XML)
library(xml2)
library(tibble)
library(tidyverse)
```

## Load Global Variables
The functions in this markdown require variables that we will define in this
chunk.

1. 'subunit_want_list' contains the agency components that we are interested in.
NOTE: This currently only applies to DHS component agencies. Change as necessary
depending on your agency of interest.

2. 'values_address_list' contains locations of data we will eventually be
extracting.

3. 'assoc_address_list' contains locations of the reference keys for each 
section.

4. 'assoc_attrs_val_list' contains locations of the reference ids for each 
section (e.g., "RD1", "PA2", etc.).

5. 'assoc_attrs_org_list' contains locations of the reference ids for an agency 
for each section.

NOTE: The addresses contained within 'values_address_list', 
'assoc_address_list', 'assoc_attrs_val_list', and 'assoc_attrs_org_list' need 
to be in the same order.

```{r}
subunit_want_list <- c("CDC", "ACF", "OIG", "HHS")

values_address_list <- c(
"/iepd:FoiaAnnualReport/foia:RequestDispositionSection/foia:RequestDisposition",
"/iepd:FoiaAnnualReport/foia:ProcessedRequestSection/foia:ProcessingStatistics",
"/iepd:FoiaAnnualReport/foia:ProcessedAppealSection/foia:ProcessingStatistics",
"/iepd:FoiaAnnualReport/foia:AppealDispositionSection/foia:AppealDisposition",
"/iepd:FoiaAnnualReport/foia:AppealResponseTimeSection/foia:ResponseTime",
"/iepd:FoiaAnnualReport/foia:ProcessedResponseTimeSection/foia:ProcessedResponseTime",
"/iepd:FoiaAnnualReport/foia:PendingPerfectedRequestsSection/foia:PendingPerfectedRequests",
"/iepd:FoiaAnnualReport/foia:SimpleResponseTimeIncrementsSection/foia:ComponentResponseTimeIncrements",
"/iepd:FoiaAnnualReport/foia:ComplexResponseTimeIncrementsSection/foia:ComponentResponseTimeIncrements",
"/iepd:FoiaAnnualReport/foia:ExpeditedResponseTimeIncrementsSection/foia:ComponentResponseTimeIncrements",
"/iepd:FoiaAnnualReport/foia:OldestPendingRequestSection/foia:OldestPendingItems/foia:OldItem/foia:OldItemPendingDaysQuantity",
"/iepd:FoiaAnnualReport/foia:ExpeditedProcessingSection/foia:ExpeditedProcessing",
"/iepd:FoiaAnnualReport/foia:PersonnelAndCostSection/foia:PersonnelAndCost",
"/iepd:FoiaAnnualReport/foia:BacklogSection/foia:Backlog",
"/iepd:FoiaAnnualReport/foia:OldestPendingConsultationSection/foia:OldestPendingItems",
"/iepd:FoiaAnnualReport/foia:OldestPendingAppealSection/foia:OldestPendingTimes",
"/iepd:FoiaAnnualReport/foia:RequestDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptions",
"/iepd:FoiaAnnualReport/foia:AppealDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptions",
"/iepd:FoiaAnnualReport/foia:FeeWaiverSection/foia:FeeWaiver", "/iepd:FoiaAnnualReport/foia:ProcessedConsultationSection/foia:ProcessingStatistics")

assoc_address_list <- c(                            "/iepd:FoiaAnnualReport/foia:RequestDispositionSection/foia:RequestDispositionOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:ProcessedRequestSection/foia:ProcessingStatisticsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:ProcessedAppealSection/foia:ProcessingStatisticsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:AppealDispositionSection/foia:AppealDispositionOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:AppealResponseTimeSection/foia:ResponseTimeOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:ProcessedResponseTimeSection/foia:ProcessedResponseTimeOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:PendingPerfectedRequestsSection/foia:PendingPerfectedRequestsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:SimpleResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:ComplexResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:ExpeditedResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:OldestPendingRequestSection/foia:OldestPendingItemsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:ExpeditedProcessingSection/foia:ExpeditedProcessingOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:PersonnelAndCostSection/foia:PersonnelAndCostOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:BacklogSection/foia:BacklogOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:OldestPendingConsultationSection/foia:OldestPendingItemsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:OldestPendingAppealSection/foia:OldestPendingItemsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:RequestDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptionsOrganizationAssociation", "/iepd:FoiaAnnualReport/foia:AppealDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptionsOrganizationAssociation",
"/iepd:FoiaAnnualReport/foia:ProcessedConsultationSection/foia:ProcessingStatisticsOrganizationAssociation")

assoc_attrs_val_list <- c(                                 "/iepd:FoiaAnnualReport/foia:RequestDispositionSection/foia:RequestDispositionOrganizationAssociation/foia:ComponentDataReference/@s:ref",         "/iepd:FoiaAnnualReport/foia:ProcessedRequestSection/foia:ProcessingStatisticsOrganizationAssociation/foia:ComponentDataReference/@s:ref",
"/iepd:FoiaAnnualReport/foia:ProcessedAppealSection/foia:ProcessingStatisticsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:AppealDispositionSection/foia:AppealDispositionOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:AppealResponseTimeSection/foia:ResponseTimeOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ProcessedResponseTimeSection/foia:ProcessedResponseTimeOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:PendingPerfectedRequestsSection/foia:PendingPerfectedRequestsOrganizationAssociation/foia:ComponentDataReference/@s:ref",
"/iepd:FoiaAnnualReport/foia:SimpleResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation/foia:ComponentDataReference/@s:ref",
"/iepd:FoiaAnnualReport/foia:ComplexResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ExpeditedResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:OldestPendingRequestSection/foia:OldestPendingItemsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ExpeditedProcessingSection/foia:ExpeditedProcessingOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:PersonnelAndCostSection/foia:PersonnelAndCostOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:BacklogSection/foia:BacklogOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:OldestPendingConsultationSection/foia:OldestPendingItemsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:OldestPendingAppealSection/foia:OldestPendingItemsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:RequestDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptionsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:AppealDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptionsOrganizationAssociation/foia:ComponentDataReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ProcessedConsultationSection/foia:ProcessingStatisticsOrganizationAssociation/foia:ComponentDataReference/@s:ref")

assoc_attrs_org_list <- c(                                  "/iepd:FoiaAnnualReport/foia:RequestDispositionSection/foia:RequestDispositionOrganizationAssociation/nc:OrganizationReference/@s:ref",                                         "/iepd:FoiaAnnualReport/foia:ProcessedRequestSection/foia:ProcessingStatisticsOrganizationAssociation/nc:OrganizationReference/@s:ref",                                         "/iepd:FoiaAnnualReport/foia:ProcessedAppealSection/foia:ProcessingStatisticsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:AppealDispositionSection/foia:AppealDispositionOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:AppealResponseTimeSection/foia:ResponseTimeOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ProcessedResponseTimeSection/foia:ProcessedResponseTimeOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:PendingPerfectedRequestsSection/foia:PendingPerfectedRequestsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:SimpleResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ComplexResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ExpeditedResponseTimeIncrementsSection/foia:ResponseTimeIncrementsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:OldestPendingRequestSection/foia:OldestPendingItemsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ExpeditedProcessingSection/foia:ExpeditedProcessingOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:PersonnelAndCostSection/foia:PersonnelAndCostOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:BacklogSection/foia:BacklogOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:OldestPendingConsultationSection/foia:OldestPendingItemsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:OldestPendingAppealSection/foia:OldestPendingItemsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:RequestDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptionsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:AppealDispositionAppliedExemptionsSection/foia:ComponentAppliedExemptionsOrganizationAssociation/nc:OrganizationReference/@s:ref", "/iepd:FoiaAnnualReport/foia:ProcessedConsultationSection/foia:ProcessingStatisticsOrganizationAssociation/nc:OrganizationReference/@s:ref")
```

## Function: ORG_want_function
Organization Component Agencies
Our first step is identifying all component agencies that we want within the
parent organization. We will then filter this dataframe to include only the 
component agencies we are interested in (e.g., for DHS we want CBP, CRCL, ICE,
OIG, PRIV, and USCIS). 

The function below will create a dataframe containing all component agencies and 
their abbreviations, as well as the reference ids (s:id -- e.g., ORG1, ORG2, 
etc.). The reference ids will be important for identifying relevant sections of 
the data (e.g., which ORG corresponds to which RD section). We will then filter 
based on the agencies in subunit_want.

The end product of this chunk is ORG_want and will act as the key for filtering
data from the XML file (e.g., ProcessedRequestSection, ProcessedAppealSection, 
etc.) 

```{r}
ORG_want_function <- function(xml_top_node,
                              xml_parsed,
                              subunit_want) {
  
  this_ORG_x <- xmlValue(getNodeSet(xml_top_node,
                        "/iepd:FoiaAnnualReport/nc:Organization/nc:OrganizationAbbreviationText"))
  
 # this_year_x <- xmlValue(getNodeSet(xml_top_node,
  #                    "/iepd:FoiaAnnualReport/foia:DocumentFiscalYearDate"))
  
  nodes_ORG_x <- getNodeSet(xml_top_node, 
                        "/iepd:FoiaAnnualReport/nc:Organization/nc:OrganizationSubUnit")
  
  df_ORG_x <- xmlToDataFrame(nodes = nodes_ORG_x, stringsAsFactors = FALSE)

  df_ORG_x$ParentOrganization <- this_ORG_x
  
 # df_ORG_x$FY <- this_year_x

  ORG_attrs_x <- xpathSApply(xml_parsed, 
                         "/iepd:FoiaAnnualReport/nc:Organization/nc:OrganizationSubUnit/@s:id")
  
  df_ORG_x$OrganizationReference <- ORG_attrs_x
  
  df_ORG_want_x <- filter(df_ORG_x, OrganizationAbbreviationText %in% subunit_want)

  return(df_ORG_want_x)
  
}

```

## Function: df_assoc_want_function 
Reference ID Key
Next, we have to make sure that the reference ids (s:id - e.g., ORG1, ORG2, 
etc.) of each agency match the reference ids (s:id - e.g., RD1, RD2, etc.) of a 
particular section (e.g., Request Disposition section in this case). 
Thankfully, there is a key provided in each section of the XML.

What we will do in the function below is find the appropriate reference id for 
a section of interest (e.g., Request Disposition). Each section contains a key
that identifies which component data reference (e.g., RD1) corresponds to a 
particular organization (e.g., ORG1). 

The for-loop is important because we use it to add values addresses to the list 
so that we can use it when extracting data.

The complete key for a section is included in the df_assoc_x dataframe. We 
will then filter this key to include only those subunit agencies we are 
interested in, so the function will return df_assoc_want_x.
```{r}
df_assoc_want_function <- function(ORG_want,
                                   values_addresses,
                                   assoc_addresses, 
                                   assoc_attrs_orgs, 
                                   assoc_attrs_vals,
                                   xml_parsed,
                                   xml_top_node) {
  df_assoc_x = c()
  
  for (i in 1:length(assoc_addresses)){
  node_assoc_sub_x <- getNodeSet(xml_top_node, assoc_addresses[i])
  df_assoc_sub_x <- xmlToDataFrame(nodes = node_assoc_sub_x, stringsAsFactors = FALSE)
  assoc_attrs_val_sub_x <- xpathSApply(xml_parsed, assoc_attrs_vals[i])
  assoc_attrs_org_sub_x <- xpathSApply(xml_parsed, assoc_attrs_orgs[i])
  df_assoc_sub_x[ , 1] <-  assoc_attrs_val_sub_x
  df_assoc_sub_x[ , 2] <- assoc_attrs_org_sub_x
  df_assoc_sub_x$values_address <- values_addresses[i]
  
  df_assoc_x <- bind_rows(df_assoc_x, df_assoc_sub_x)
}

  df_assoc_want_x <- filter(df_assoc_x, OrganizationReference %in%
                              ORG_want$OrganizationReference)

  df_assoc_want_x$Section <- str_extract(df_assoc_want_x$values_address, '(?<=:)\\w*$')
  
  return(df_assoc_want_x)
  
}

```

## Function: section_values_function
Extract Data for a Section
Ok, now we can start extracting data from a section (e.g., Request Disposition). 

This function will go through the a section and extract values associated with
the desired ID-X (e.g., RD1) attributes, which corresponds to ORG-X (ORG1), 
which corresponds to a given agency in the parent data set (e.g., CBP in the
DHS data set). We will create a 'for' loop that completes this process for each 
subunit of interest. We will also append the section id (RD-X) and organization
id (ORG-X) to this data frame.

This function will return section_values_full_x.
```{r}
section_values_function <- function(ORG_want,
                                    assoc_want,
                                    xml_top_node,
                                    values_address) {
  section_values_year_x <- data.frame()
  
  for (i in 1:nrow(assoc_want)){

  node_section_subset_x <- getNodeSet(xml_top_node, paste0(values_address,
                                                           "[@s:id ='", 
                                                           assoc_want$ComponentDataReference[i],
                                                    "']"))
  
  extract_section_subunit_subset_x = xmlSApply(node_section_subset_x, 
                                               function(x) xmlSApply(x, xmlValue))
  section_subunit_values_subset_x = data.frame(t(extract_section_subunit_subset_x), 
                                               row.names = NULL)
  
  section_subunit_values_subset_x$ComponentDataReference <- assoc_want$ComponentDataReference[i]
  section_subunit_values_subset_x$OrganizationReference <- assoc_want$OrganizationReference[i]

  section_values_year_x <- bind_rows(section_values_year_x, section_subunit_values_subset_x)
  }
  
  section_values_year_x <- section_values_year_x %>% left_join(ORG_want,
                                                               by = "OrganizationReference")
  
  section_values_year_x <- section_values_year_x %>% left_join(assoc_want,
                                                               by = "ComponentDataReference") %>% 
                                                     select(-ends_with(".y"))
  
  if ("t.extract_section_subunit_subset_x." %in% colnames(section_values_year_x)) {
      section_values_year_x <- select(section_values_year_x, -"t.extract_section_subunit_subset_x.")
    }
  
  return(section_values_year_x)
}  

```

## Extract and save data from XML files in a folder
In the previous chunks we created global variables and functions that we will
use here. First, we start by identifying all file names within a folder. For
now, let's stick with files pertaining to all years for a single agency 
(e.g., DHS) in a folder. This will make it easier to troubleshoot any potential 
issues.

The for-loop will call all the functions we defined earlier and the end product
will be section_values_all, containing all values for all sections for all 
years for the agency of interest. This data frame will then be saved in the 
working directory. Example file name: DHS_section_values_all.Rda.
```{r}
filenames = dir(path = "./FOIA Sample")
# Create an empty list to store the data frames for each item in values_address_list
foia_data <- list()

for (i in filenames) {
  # Import and parse the XML file so we have a workable R format.
  xml_1 <- xmlParse(paste0("FOIA Sample/", i))
  
  # Get the top-level node in the XML document we just parsed. This will allow
  # us to more easily locate nodes of interest.
  xml_top <- xmlRoot(xml_1)
  
  df_ORG_want <- ORG_want_function(
    xml_top_node = xml_top,
    xml_parsed = xml_1,
    subunit_want = subunit_want_list
  )
  
  df_assoc_want <- df_assoc_want_function(
    ORG_want = df_ORG_want,
    values_addresses = values_address_list,
    assoc_addresses = assoc_address_list,
    assoc_attrs_orgs = assoc_attrs_org_list,
    assoc_attrs_vals = assoc_attrs_val_list,
    xml_parsed = xml_1,
    xml_top_node = xml_top
  )
  
  # Loop through each item in values_address_list
  for (address in values_address_list) {
    # Filter based on section name.
    df_assoc_want_filter <- filter(df_assoc_want,
                                   Section == str_extract(address, '(?<=:)\\w*$'))
    
    section_values_data <-
      section_values_function(
        ORG_want = df_ORG_want,
        assoc_want = df_assoc_want_filter,
        xml_top_node = xml_top,
        values_address = address
      )
    
    # Add filename and FY to section_values_data
    section_values_data$filename <- i
    section_values_data$FY <- str_extract(i, '[:digit:]+')
    
    # Create a unique data frame name based on the section.
    section_data_name <- str_extract(address, 'foia:([^/]+)')
    section_data_name <- str_remove(section_data_name, 'foia:')
    
    # If section_data_name already exists in foia_data and the number
    # of columns match, append to it. Before appending, make sure that the order
    # of the column names match.
    if (section_data_name %in% names(foia_data) &
        length(colnames(section_values_data)) == length(colnames(foia_data[[section_data_name]]))) {
      col_order <- colnames(foia_data[[section_data_name]])
      
      section_values_data <- section_values_data[col_order]
      
      foia_data[[section_data_name]] <-
        rbind(foia_data[[section_data_name]], section_values_data)
      
    } else if (section_data_name %in% names(foia_data) &
               length(colnames(section_values_data)) != length(colnames(foia_data[[section_data_name]]))) {
      # Check for missing columns between data frames and add the missing columns.
      missing_columns <-
        dplyr::setdiff(colnames(foia_data[[section_data_name]]),
                       colnames(section_values_data))
      for (column in missing_columns) {
        section_values_data[[column]] <- NA
      }
      
      # Make sure columns are in the same order
      col_order <- colnames(foia_data[[section_data_name]])
      
      section_values_data <- section_values_data[col_order]
      
      # Append to foia_data
      foia_data[[section_data_name]] <-
        rbind(foia_data[[section_data_name]], section_values_data)
      
    } else {
      # If section_data_name doesn't exist, create a new entry in the list
      foia_data[[section_data_name]] <- section_values_data
    }
  } # To access the data in a section you can use foia_data$SectionName (e.g.,
}   # foia_data$RequestDisposition) or foia_data[["SectionName"]].


```
