---
title: "data comparison"
author: "John Lee"
date: "2025-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

## Data Comparison

Compare original DHS foia data to newly created one.

The document assumes that the original `DHS_data.Rda` and `DHS_PPR_data.Rda` (which were created using the original data mining functions) are in the `../../code/data/rda/` directory. The new `DHS_data.Rda` is assumed to be in the `../data/` directory.

```{r}
load("../../code/data/rda/DHS_data.Rda")
orig_foia <- foia_data

load("../data/DHS_data.Rda")
new_foia <- foia_data

load("../../code/data/rda/DHS_PPR_data.Rda")
orig_foia_ppr <- foia_data
```

### RequestDispositionSection

```{r}
all.equal(orig_foia$RequestDispositionSection, new_foia$RequestDispositionSection)
```

### ProcessedRequestSection

Notice that the original data scraped data from non existent ComponentDataReference due to a bug in the code.

```{r eval=FALSE}
View(orig_foia$ProcessedRequestSection)
View(new_foia$ProcessedRequestSection)
```

```{r}
all.equal(
  orig_foia$ProcessedRequestSection %>%
    filter(!if_all(starts_with("Process"), ~ is.na(.x))),
  new_foia$ProcessedRequestSection %>%
    filter(FY != 23)
)
```

### OldestPendingAppealSection

```{r eval=FALSE}
View(orig_foia$OldestPendingAppealSection)
View(new_foia$OldestPendingAppealSection)
```

The original foia dataset is missing all data

### OldestPendingRequestSection

```{r eval=FALSE}
View(orig_foia$OldestPendingRequestSection)
View(new_foia$OldestPendingRequestSection)
```

Same here

### PendingPerfectedRequestsSection

Important thing to note is that if the data comes from nested nodes, then the node names are separated with a period in the new foia data. This wasn't the case for old foia data.

Columns:

```{r}
print(colnames(orig_foia_ppr$PendingPerfectedRequestsSection))
print(colnames(new_foia$PendingPerfectedRequestsSection))
```

```{r eval=FALSE}
View(orig_foia_ppr$PendingPerfectedRequestsSection)
View(new_foia$PendingPerfectedRequestsSection)
```

```{r}
tmp <- new_foia$PendingPerfectedRequestsSection %>%
  filter(FY != 23) %>%
  # all "_new" to all columns
  rename_with(~ paste0(.x, "_new")) %>%
  rename_with(~ str_replace(.x, "ComplexPendingRequestStatistics_", "ComplexPendingRequestStatistics")) %>%
  rename_with(~ str_replace(.x, "SimplePendingRequestStatistics_", "SimplePendingRequestStatistics")) %>%
  rename_with(~ str_replace(.x, "ExpeditedPendingRequestStatistics_", "ExpeditedPendingRequestStatistics")) %>%
  bind_cols(orig_foia_ppr$PendingPerfectedRequestsSection) %>%
  select(order(colnames(.)))

for (i in 1:(ncol(tmp) / 2)) {
  print(all.equal(tmp[[(i-1)*2 + 1]], tmp[[i*2]]))
}
```

The values across the corresponding columns match exactly.
